### 2024-09-16 texts/paper_closing.tex
minor polishing, update tasks, send email, update sups

% Created 2024-09-11 wo 11:36
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

\section*{It's so over}

@@

In recent years, private collector-run museums have seen an unprecedented proliferation, with hundreds of organizations founded since the beginning of the millenium \parencite{Velthuis_etal_2023_boom,LarrysList_2015_report}.
These museums, run by individual collectors, constitute an addition to established organizational forms of art provision such as museums operated and financed by the state, foundations or associations.
While many aspects of this new organizational form has attracted attention, such as their emergence, relations to other institutions or impact on artist consecration (cf. \cite{Kolbe_etal_2022_privatemuseum} for a literature review), the process of closure has received considerably less attention
Furthermore, the focus that has been given to questions of sustainability has been primarily conducted through case studies; beyond the exception of \textcite{Velthuis_Gera_2024_fragility} quantitative studies have been absent.
In this paper, I address this gap by investigating the closure event of members of this new organizational form.
Analysis which factors are associated with museum closures provides insight not only into the event of closing, but also into the ongoing processes of a private museums, such as dependency relationships.
Furthermore, analyzing which museums survive allows to make informed speculations on the future of this recently emerged organizational form.

While a recent study by \textcite{Velthuis_Gera_2024_fragility} has investigated closing reasons of private museums, I extend/complement this study by using a more elaborate statistical model and additional data sources.
In particular, I compare private museum closing rates to those of other museum populations, and analyze the risk associated with variables on the level of the founder, museum and environment. 
Taken together, the results show that (at least so far) little evidence exists neither for private museums being overly dependend on their founder, nor having overall higher mortality rates than other non-profit museums.
While little evidence points toward private museums being unvoluntarily at risk due to their low diversification, private museums may close due to environmental pressures, which exhibit a complex relationship which for most museums consists of competition from other private museums as well as, unexpectedly, large audiences.
Finally, some limited evidence points towards a temporary intention by their founders from the very beginning as a potential reason for their closure.

@@

Due to the founders central role, private museums have been argued to be highly dependent on their founder for their continuing existence, as museums generally cannot cover expenses via revenues from ticket prices, museum shops and cafes:
The \textcite{IMLS_2008_funding} found that for non-profit American art museums, less than half (48.1\%) the revenues come from earned income, leaving substantial gaps in the budgets that are filled by a mix of private donations (23.0\%), investment (18.5\%) and government support (10.4\%) (p.27).
Outside of the United States, the reliance on revenue streams beyond operations is even larger:
German museums in the 1990s covered on average only around 30\% of their budget by their income \parencite{Martin_1993_museen}\footnote{Own calculations based on table p.233}, with private museums requiring less support as operations provide around 47\% of their budget (ibid.), and even the more market-oriented NPOs in the cultural sector draw around 50\% of their funding from public support \parencite[p.82]{Zimmer_Priller_2007_gemeinnuetzig}.
On a European level, only 12\% of museum incomes are generated by entry fees, whereas 69\% originate from subsidies \parencite{EGMUS_2024_complete.}\footnote{Own calculations from on EGMUS data, based on all countries with income data from 2015 onwards.}

@@

A central reason for financial vulnerability has been argued to be the crucial reliance on the single founder, who, as the iniators of their private museums, likely have to contribute a large part of their institution's non-operating income \parencite{Frey_Meier_2002_beyeler}:
\textcite{Adam_2021_rise} argues that "ultimately, many of these failures demonstrate the fragility of spaces that rely on a single founder, whose motivations and financial stability may change, or who may not have realised the difficulties inherent in establishing their own art space" (p.82).
\textcite{Velthuis_Gera_2024_fragility} similarly argue that "because of their funding models and reliance on a sole founder, [private museums] are inherently fragile organizations" (p.1).
\textcite{Bechtler_Imhof_2018_future} also characterize "the dependency on single individuals only" as one of main reasons that "make these private museums fragile" (p.53).
\cite{StylianouLambert_etal_2014_museums} also characterize many private institutions as "small individual museums that depend too much on their creators to guarantee their sustainability" (p.582).
Beyond a direct financial component, private museums have been also been characterized as fragile since they can only rely on their founder for "taste", "vision and drive" \parencite[p.77]{Adam_2021_rise}, "passion" \parencite[p.234]{Walker_2019_collector} or "skills and knowledge" \parencite[p.580]{StylianouLambert_etal_2014_museums}.
The link between diversification and mortality, i.e. higher survival chances for more diversified organizations, has been found for non-profit organizations more generally \parencite{Fernandez_2007_dissolution,Bielefeld_1994_survival,Hager_2001_vulnerability,Lu_Shon_Zhang_2019_dissolution}; it is thus plausible to apply to museums as well.

Lack of diversification is then argued to pose a problem in particular in the case of the founder's death:
\textcite{Walker_2019_collector} argues that "the death of the original founder and creator can [\ldots{}] place the future of private museums and collections in jeopardy" as "seldom do heirs share a similar passion or wish to take on the financial burden of maintaining private museums indefinitely" (p.234). 
\textcite{Bechtler_Imhof_2018_future} argue that in order to ensure museum longevity, "flexible, easily adapted designs for private museums" are necessary "to reduce the financial burden on their heirs" (p.188).
\textcite{Adam_2021_rise} argues that succession introduces strains in financial aspects ("Collectors’ successors may not look kindly on what may well be a money pit when [\ldots{}] they will have to pick up the bills", p.77), which also concerns an aesthetic dimension ("the nature of contemporary art is to be fresh and new, but what is ‘fresh and new’ for the founder might seem ‘old hat’ to their heirs", p.76).
\textcite{Velthuis_Gera_2024_fragility} similarly argue that "the founder’s death will almost inevitably have a direct impact on the financial health of the museum" (p.10).
\cite{StylianouLambert_etal_2014_museums}, after discussing a case of a private museum whose owner "does not believe that anyone else could love the museum as much as she does", argue that this "generates questions regarding the cultural sustainability of small private museums in cases where their initiators are not able to care for them any longer" (p.580).

@@

A newly constructed database, assembled from a wide range of sources, is used to measure the survival of private museums (cf. \textcite{Velthuis_etal_2023_boom} for details).
Currently, the database includes 548 museums, of which 451 are currently open, 72 have closed and 25 have transformed into other organizational forms.\footnote{The 25 institutions which were private museums at some point and have since been transformed into other organizational forms are not included in this analysis and discussed in the limitations.}
Using this database, in particular year of opening and, if available, closing, it is possible to reconstruct the life course of the each institution with museum-year as the unit of analysis.
While the database is not limited to private museums of a particular time period, due to the recency of the private museum boom in the last two decades 80\% of the museums included in it have opened after 2000.
As the databse was constructed retrospectively from 2020 to 2022, I focus here on closing events from 2000 to 2021 to limit the possibility that some institutions were missed, as such risks are likely higher for earlier time periods.
Furthermore, a small number of museums are excluded, in particular 2 museums for which the closing date could not be found, 15 museums which opened in 2021 and later (as survival analyses requires positive survival times), 1 museum the opening year of which could not be determined, as well as 2 which closed before 2000 thereby fall outside the period under investigation.
This leaves 503 private museums at risk which have accumulated 6096 museum-years and 68 closures.
This data allows to conduct survival analysis (also known as event history analysis), the most common statistical technique to investigate hazard rates and identify mortality-assocated covariates \parencite{Moore_2015_survival,Allison_2014_event}.

@@

To ensure that inclusion of the word "museum" corresponds to the naming convention of the museum population, the names of the "Museums of the World" (MOW) database \parencite{deGruyter_2021_MOW}, a database containing over 55.000 museums, are analyzed.
In this database, 71\% of museum names contain the term "museum" (or its equivalent in other languages), which indicates a substantial convention by founders to include the term museum in the name of their institution\footnote{While it would also be possible to construct inclusion in the MOW as an indicator of legitimacy, as it represents recognition of the organization as a museum by experts \parencite{Zuckerman_1999_illegitimacy}, preliminary investigations of the database indicated that it has been  updated only sparsely in recent years; thereby inclusion in it (or the lack therefore) does not constitute an accurate measurement of the extent of an institutions' recognition as a museum}.
\paragraph*{Reputation}

@@

While I did not find a systematic literature review or meta-analysis, a number of individual studies had investigated similar populations based on IRS data (see table \ref{tbl:litreview})\footnote{IRS data is not without limitations, such as only being required from larger organizations, misreporting, and the difficulty of inferring closing from non-filing. Nevertheless, it constitutes the most fine-grained data source to investigate non-profit survival}.
None of these focus on art museums exclusively (\textcite{Hager_2001_vulnerability} focuses on Art NPOs, \textcite{Bowen_etal_1994_charitable} and \textcite{Gordon_etal_2013_insolvency} on NPOs generally), but report survival estimates by organizational form, including museums.
\textcite{Bowen_etal_1994_charitable} investigates the closing of 501(c)(3) museums in the 1981-1991 time period and, based on 271 (9.8\%) closures of a population of 2755, calculate average annual exit rates as 1.1\% (p.103).
\textcite{Gordon_etal_2013_insolvency} observes 35 closures over the years of 2000-2003 with 5,167 museum-years (p.368), resulting in an average annual closing rate of 0.68\% (p.377).
Some of this difference may be explained by higher shares of smaller organizations in \textcite{Bowen_etal_1994_charitable}, where museums of any size are included, whereas \textcite{Gordon_etal_2013_insolvency} excluded NPOs with operational revenues below 150,000 USD, as smaller non-profits are at higher risk of closing \parencite{Harrison_Laincz_2008_nonprofit,Helmig_Ingerfurth_Pinz_2013_nonprofit}.

@@

Table \ref{tbl:t_reg_coxph_deathcfg} shows coefficients of both periods with different cut-off points.
While the coefficient in the short-term period is consistently higher than both the overall death coefficient and the long-term period coefficients, so far the evidence is inconclusive as only few cases are statistically significant\footnote{P-values are also not adjusted for multiple tests}.
Furthermore, such a shock passes quickly relative to the median life-time of museums of 52 years, thus "giving" the founder death little time to lead to many museum closures. 
Considering the numbers underlying the death effect estimates illustrates the uncertainty of the determning the effect of the death of the founder on aclosing: 
While 47 private museums have seen the death of their founders, only 3 of these have closed in the following 5 years; given such low numbers it cannot be ruled out that the statistical significance in a few configurations of founder death time results from random fluctuations rather than an underlying pattern.

@@

At the local level, the number of private museum and population numbers predict survival chances as the main effects of population and number of private museums as well as its interaction are significant.
The particular pattern however is complex:
As both main terms are positive and the interaction negative, at low values increases in either variable are associated with increases in mortality (figure \ref{fig:p_condmef}).
Thus, for population values of 0 to 2.97 million people, an increase in the number of private museums corresponds to an increase in mortality (bottom half of figure \ref{fig:p_pred_heatmap}), which provides support for hypothesis 3 according to which competition leads to museum closure, in particular as this range includes 5056 museums years, or 83\% of the dataset. 
Yet at the same time, at low private museums numbers (0 to 3 private museums), increases in population also increase mortality; this range also includes a large number of museum-years (4960 museum years; 81\% of the data).
Such a relationship is unexpected insofar as it is unclear why a larger potential audience would be associated with a higher risk of closure.\footnote{Alternative approaches of modeling competition are discussed in the supplementary materials, but cannot answer this question.}
A possible explanation might be that collectors in less densely populated areas are more committed than those in less populated areas precisely because they know they cannot count on visitors to contribute substantially.
Alternatively, areas with high population and low private museum numbers might more generally lack cultural infrastructure; which one could see supported by closures in Jakarta, Istanbul and Mexico City, but less so by closings in Paris, Tokyo or Madrid.
However, low private museums indicating lack of cultural infrastructure might also explain the decrease in mortality associated with increases in private museum numbers at high population values (figure \ref{fig:p_pred_heatmap}, upper half), as such pattern does not correspond to that predicted by competition.
More in line with expectation is however the association between increases of population numbers and lower mortality at higher numbers of private museums (right half), as it indicates that audiences can facilitate museum survival.

@@

Museums which include the term "collections" in their name are the least likely to close (0.63 times the rate of museums with the name "museum", while those with the term foundation experience 23\% elevated closure rates (but not significantly higher).
However, institutions which include neither of these are the most likely to close, and have mortality rates almost three times as high (198\%) than institutions including the term "museum".
The usage of unconventional names might reflect adherence to museum standards which makes museums more understandable and familiar to third parties (e.g. audiences, other museums, companies or the government) and thereby facilitates interactions which are beneficial for long-term survival (e.g. attracting visitors, cooperations, sponsorships or subsidies).
However, as founders can choose the name of their museum arbitrarily, they might choose a name without reference to organizational forms associated with perpetuity precisely because their intention is not an institution enduring in perpetuity but rather a more temporally bounded project.
Some support for the latter interpretaton is given by the fact that two of the three museums for which \textcite{Velthuis_Gera_2024_fragility} identify a temporal intention as the closing reason do not self-identify as museum, collection or foundation.
A temporal intentions are however the exceptions among closing reasons identified by \textcite{Velthuis_Gera_2024_fragility} (p.6), and as similarly only a small fraction of private museum chooses such a name (140 museums, or 28\%), the overall proportion of these types of voluntary closures compared to all closures is presumably limited.
\subsection*{Discussion}

Previously the dozens of closures have been seen as indicator of their vulnerability and short-livedness \parencite{Adam_2020_close}, however this analysis shows the importance studying closing not in isolaton but in comparison to both non-closures and closures of comparable institutions.
While the booming numbers of private museums themselves have received substantial attention, less consideration has been given to the fact that large populations can generate dozens of closures without requiring high closing rates, especially when cloures are considered over a span of several years or decades.
In other words, the sheer size to which the private museum population has grown over the last decades has made it all but inevitable that some of them would close, and do so after only a few year.
Based on the current analysis, private museums will continue to close in the years to come; at hazard rates of around 1\% on average a handful of closures every year are to be expected.

Comparing private museums closures to closures of other organizational form advances our understanding of the sustainability of this new organizational form.
While private museums do close, the comparison to previous studies of museum closing rates provides no evidence that dependence on single founders constitutes a less stable organizational form than reliance on less centralized foundations, associations or corporate sponsorship.
Multiple reasons might be the at play that mitigate this risk of low diversification.
First, it might take more commitment to open a museum as an individual or couple compared to less involved art patronage of being a board member in existing institutions or donating works, which would result in private museums founders being more motivated to secure the existence of their institutions.
Furthermore, the centrality of the founder might leave little room to question who is responsible to provide additional funds in times of financial difficulties, whereas in organizations without a central figure each member may feel less responsible.
It might also be such commitment that leads founder to plan the future of their museum to such an extent that their death is not significantly associated with unvoluntary closing.

While so far little evidence points toward low diversification, i.e. reliance on the founder, as a key vulnerability of private museums to unvoluntary closures, the centrality of the founder may be more pronounced in voluntary closures. 
Indicated by unconventional naming choice, a subset of founders might from the very beginning envision the museum as temporary.
On the other hand, higher closing rates for unconventionally named museums might reflect absence of recognition and support by third parties in the environment.
A dimensions where the influence of the environment can however be ascertained with greater certainty is the role of audiences and other private museums.
While the overall relationship is complex, for the majority of private museums increases in both variables are associated with higher closing rates.
Whereas it is plausible that the association between higher institution counts and closing rates reflects competitive pressures, it is less obvious what may lay behind observing a positive association behind larger population and closing risk. 
\subsection*{Conclusion}

\subsubsection*{Limitations}

However, it is necessary to consider that so far only a small fraction (47 museums, 9.3\%) of museums has experienced the death of their founder, one of the primary reasons hypothesized to lead to the private museum closure. 
While so far these museums have not seen substantially elevated closing risks, these estimates will become more precise as more founders pass away.
As the currently large confidence interval (figure \ref{fig:p_surv_death}) for museums whose founders have died allows a wide range of survival chances, it is entirely possible that previous scholars have indeed correctly identified a key dependency (the current absence of evidence of it should not be taken as evidence of absence) and that therefore later analyses will a significantly association between founder death and closing risk as well as a higher average closing risk for private museums.
At the same time, it is also possible that private museums are generally less dependent on a single person (or couple) than has been assumed, and have in fact become institutionalized to an extent that most heirs of founders feel as motivated to continue the original founders' efforts as these founders themselves.

@@

The primary mechanism that has been argued to lead to museum closure is financials, and while this study measures it indirectly, primarily through indicators on the level of the founder, museum and environment, the financial situation of a museum is not used directly.
The primary reason for this is data availability.
While centralized non-profit financial information is available for American non-profits through the IRS/NCCS files, the much larger private museum population in the rest of the world is not subject to similar reporting requirements.
The effort to compile such information on a global scale ranges from extremely challenging (harmonizing different reporting standards, language barriers) to outright impossible for institutions who do not even (have to) report public financial information.
Similarly, other museum characteristics, such as as museum size, facilities, as well as marketing and outreach activities might be related to long-term survival prospect\footnote{While the original data collection included the assessment of various museum characteristics, such as floor and collection size, facilities, and cooperations, it became clear during data analysis that the data collected was unsuitable to construct accurate indicators from these findings due to difficulty in assessing them in many cases.
Especially for closed museums data collection proved difficult, as in many cases websites had been shut down, and while in many cases snapshots of the main website were available in the Wayback machine, less central pages (such as those with information about facilities and activities), were archived much less frequently, and sometimes inaccessible despite being archived due to changes in webtechnology (such as the retirement of Adobe flash).}
As such, private museums remain relatively intransparent, which does not prevent, but to some extent limits the identification of risk factors.

>>>>

% Created 2024-09-11 wo 15:33
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

\section*{Stayin' Alive: The unexpected longevity of founder-centered private art museums}

@@

In recent years, private collector-run museums have seen an unprecedented proliferation, with hundreds of organizations founded since the beginning of the millenium \parencite{Velthuis_etal_2023_boom,LarrysList_2015_report}.
These museums, run by individual collectors, constitute an addition to established organizational forms of art provision such as museums operated and financed by the state, foundations or associations.
While many aspects of this new organizational form has attracted attention, such as their emergence, relations to other institutions or impact on artist consecration (cf. \cite{Kolbe_etal_2022_privatemuseum} for a literature review), the process of closure has received considerably less attention. 
Furthermore, the focus that has been given to questions of sustainability has been primarily conducted through case studies; quantitative studies have been mostly absent.
In this paper, I address this gap by investigating the closure event of members of this new organizational form.
Analysis which factors are associated with museum closures provides insight not only into the event of closing, but also into the ongoing processes of a private museums, such as dependency relationships.
Furthermore, analyzing which museums survive allows to make informed speculations on the future of this recently emerged organizational form.

While a recent study by \textcite{Velthuis_Gera_2024_fragility} has investigated closing reasons of private museums, I extend/complement this study by using a more elaborate statistical model and additional data sources.
In particular, I compare private museum closing rates to those of other museum populations, and analyze the risk associated with variables on the level of the founder, museum and environment. 
Taken together, the results show that (at least so far) little evidence exists neither for private museums being overly dependend on their founder, nor having overall higher mortality rates than other non-profit museums.
While little evidence points toward private museums being unvoluntarily at risk due to their low diversification, private museums may close due to environmental pressures, which exhibit a complex relationship which for most museums consists of competition from other private museums as well as, unexpectedly, large audiences.
Finally, private museums with unconventional names (in particular, without the term "museum" in the name) face a higher closing risk, which may indicate either a temporary intention by their founders from the very beginning and/or limited recognition by third parties as potential closing reasons.

@@

Due to the founders central role, private museums have been argued to be highly dependent on their founder for their continuing existence, as museums generally cannot cover expenses via revenues from ticket prices, museum shops and cafes:
The \textcite{IMLS_2008_funding} found that for non-profit American art museums, less than half (48.1\%) the revenues come from earned income, leaving substantial gaps in the budgets that are filled by a mix of private donations (23.0\%), investment (18.5\%) and government support (10.4\%) (p.27).
Outside of the United States, the reliance on revenue streams beyond operations is even larger:
German museums in the 1990s covered on average only around 30\% of their budget by their income \parencite{Martin_1993_museen}\footnote{Own calculations based on table p.233}, with private museums however requiring less support as operations provide around 47\% of their budget (ibid.), and even the more market-oriented NPOs in the cultural sector draw around 50\% of their funding from public support \parencite[p.82]{Zimmer_Priller_2007_gemeinnuetzig}.
On a European level, only 12\% of museum incomes are generated by entry fees, whereas 69\% originate from subsidies \parencite{EGMUS_2024_complete}.\footnote{Own calculations from on EGMUS data, based on all countries with income data from 2015 onwards.}

@@

A central reason for financial vulnerability has been argued to be the crucial reliance on the single founder, who, as the iniators of their private museums, likely have to contribute a large part of their institution's non-operating income \parencite{Frey_Meier_2002_beyeler}:
\textcite{Adam_2021_rise} argues that "ultimately, many of these failures demonstrate the fragility of spaces that rely on a single founder, whose motivations and financial stability may change, or who may not have realised the difficulties inherent in establishing their own art space" (p.82).
\textcite{Velthuis_Gera_2024_fragility} similarly argue that "because of their funding models and reliance on a sole founder, [private museums] are inherently fragile organizations" (p.1).
\textcite{Bechtler_Imhof_2018_future} also characterize "the dependency on single individuals only" as one of main reasons that "make these private museums fragile" (p.53).
\textcite{StylianouLambert_etal_2014_museums} also characterize many private institutions as "small individual museums that depend too much on their creators to guarantee their sustainability" (p.582).
Beyond a direct financial component, private museums have been also been characterized as fragile since they can only rely on their founder for "taste", "vision and drive" \parencite[p.77]{Adam_2021_rise}, "passion" \parencite[p.234]{Walker_2019_collector} or "skills and knowledge" \parencite[p.580]{StylianouLambert_etal_2014_museums}.
The link between diversification and mortality, i.e. higher survival chances for more diversified organizations, has been found for non-profit organizations more generally \parencite{Fernandez_2007_dissolution,Bielefeld_1994_survival,Hager_2001_vulnerability,Lu_Shon_Zhang_2019_dissolution}; it is thus plausible to apply to museums as well.

Lack of diversification is then argued to pose a problem in particular in the case of the founder's death:
\textcite{Walker_2019_collector} argues that "the death of the original founder and creator can [\ldots{}] place the future of private museums and collections in jeopardy" as "seldom do heirs share a similar passion or wish to take on the financial burden of maintaining private museums indefinitely" (p.234). 
\textcite{Bechtler_Imhof_2018_future} argue that in order to ensure museum longevity, "flexible, easily adapted designs for private museums" are necessary "to reduce the financial burden on their heirs" (p.188).
\textcite{Adam_2021_rise} argues that succession introduces strains in financial aspects ("Collectors’ successors may not look kindly on what may well be a money pit when [\ldots{}] they will have to pick up the bills", p.77), which also concerns an aesthetic dimension ("the nature of contemporary art is to be fresh and new, but what is ‘fresh and new’ for the founder might seem ‘old hat’ to their heirs", p.76).
\textcite{Velthuis_Gera_2024_fragility} similarly argue that "the founder’s death will almost inevitably have a direct impact on the financial health of the museum" (p.10).
\textcite{StylianouLambert_etal_2014_museums}, after discussing a case of a private museum whose owner "does not believe that anyone else could love the museum as much as she does", argue that this "generates questions regarding the cultural sustainability of small private museums in cases where their initiators are not able to care for them any longer" (p.580).

@@

A newly constructed database, assembled from a wide range of sources, is used to measure the survival of private museums (cf. \textcite{Velthuis_etal_2023_boom} for details).
Currently, the database includes 548 museums, of which 451 are currently open, 72 have closed and 25 have transformed into other organizational forms.\footnote{The 25 institutions which were private museums at some point and have since been transformed into other organizational forms are not included in this analysis and discussed in the limitations.}
Using this database, in particular year of opening and, if available, closing, it is possible to reconstruct the life course of the each institution with museum-year as the unit of analysis.
While the database is not limited to private museums of a particular time period, due to the recency of the private museum boom in the last two decades 80\% of the museums included in it have opened after 2000.
As the databse was constructed retrospectively from 2020 to 2022, I focus here on closing events from 2000 to 2021 to limit the possibility that some institutions were missed, as such risks are likely higher for earlier time periods.
Furthermore, a small number of museums are excluded, in particular 2 museums for which the closing date could not be found, 15 museums which opened in 2021 and later (as survival analyses requires positive survival times), 1 museum the opening year of which could not be determined, as well as 2 museums which closed before 2000 and thereby fall outside the period under investigation.
This leaves 503 private museums at risk which have accumulated 6096 museum-years and 68 closures.
This data allows to conduct survival analysis (also known as event history analysis), the most common statistical technique to investigate hazard rates and identify mortality-assocated covariates \parencite{Moore_2015_survival,Allison_2014_event}.

@@

To ensure that inclusion of the word "museum" corresponds to the naming convention of the museum population, the names of the "Museums of the World" (MOW) database \parencite{deGruyter_2021_MOW}, a database containing over 55,000 museums, are analyzed.
In this database, 71\% of museum names contain the term "museum" (or its equivalent in other languages), which indicates a substantial convention by founders to include the term museum in the name of their institution.\footnote{While it would also be possible to construct inclusion in the MOW as an indicator of legitimacy, as it represents recognition of the organization as a museum by experts \parencite{Zuckerman_1999_illegitimacy}, preliminary investigations of the database indicated that it has been  updated only sparsely in recent years; thereby inclusion in it (or the lack therefore) does not constitute an accurate measurement of the extent of an institutions' recognition as a museum.}
\paragraph*{Reputation}

@@

While I did not find a systematic literature review or meta-analysis, a number of individual studies had investigated similar populations based on IRS data (see table \ref{tbl:litreview}).\footnote{IRS data is not without limitations, such as only being required from larger organizations, misreporting, and the difficulty of inferring closing from non-filing. Nevertheless, it constitutes the most fine-grained data source to investigate non-profit survival}
None of these focus on art museums exclusively (\textcite{Hager_2001_vulnerability} focuses on Art NPOs, \textcite{Bowen_etal_1994_charitable} and \textcite{Gordon_etal_2013_insolvency} on NPOs generally), but report survival estimates by organizational form, including museums.
\textcite{Bowen_etal_1994_charitable} investigates the closing of 501(c)(3) museums in the 1981-1991 time period and, based on 271 (9.8\%) closures of a population of 2755, calculate average annual exit rates as 1.1\% (p.103).
\textcite{Gordon_etal_2013_insolvency} observes 35 closures over the years of 2000-2003 with 5,167 museum-years (p.368), resulting in an average annual closing rate of 0.68\% (p.377).
Some of this difference may be explained by higher shares of smaller organizations in \textcite{Bowen_etal_1994_charitable}, where museums of any size are included, whereas \textcite{Gordon_etal_2013_insolvency} excluded NPOs with operational revenues below 150,000 USD, as smaller non-profits are at higher risk of closing \parencite{Harrison_Laincz_2008_nonprofit,Helmig_Ingerfurth_Pinz_2013_nonprofit}.

@@

Table \ref{tbl:t_reg_coxph_deathcfg} shows coefficients of both periods with different cut-off points.
While the coefficient in the short-term period is consistently higher than both the overall death coefficient and the long-term period coefficients, so far the evidence is inconclusive as only few cases are statistically significant\footnote{P-values are also not adjusted for multiple tests}.
Furthermore, such a shock passes quickly relative to the predicted median life time of museums of 52 years, thus "giving" the founder death little time to lead to many museum closures. 
Considering the numbers underlying the death effect estimates illustrates the uncertainty of the determning the effect of the death of the founder on aclosing: 
While 47 private museums have seen the death of their founders, only 3 of these have closed in the following 5 years; given such low numbers it cannot be ruled out that the statistical significance in a few configurations of founder death time results from random fluctuations rather than an underlying pattern.

@@

At the local level, the number of private museum and population numbers predict survival chances as the main effects of population and number of private museums as well as its interaction are significant.
The particular pattern however is complex:
As both main terms are positive and the interaction negative, at low values increases in either variable are associated with increases in mortality (figure \ref{fig:p_condmef}).
Thus, for population values of 0 to 2.97 million people, an increase in the number of private museums corresponds to an increase in mortality (bottom half of figure \ref{fig:p_pred_heatmap}), which provides support for hypothesis 3 according to which competition leads to museum closure, in particular as this range includes 5056 museums years, or 83\% of the dataset. 
Yet at the same time, at low private museums numbers (0 to 3 private museums, left half of figure \ref{fig:p_pred_heatmap}), increases in population also increase mortality; this range also includes a large number of museum-years (4960 museum years; 81\% of the data).
Such a relationship is unexpected insofar as it is unclear why a larger potential audience would be associated with a higher risk of closure.\footnote{Alternative approaches of modeling competition are discussed in the supplementary materials, but cannot answer this question.}
A possible explanation might be that collectors in less densely populated areas are more committed than those in less populated areas precisely because they know they cannot count on visitors to contribute substantially.
Alternatively, areas with high population and low private museum numbers might more generally lack cultural infrastructure; which one could see supported by closures in Jakarta, Istanbul and Mexico City, but less so by closings in Paris, Tokyo or Madrid.
However, low private museums numbers indicating lack of cultural infrastructure might also explain the decrease in mortality associated with increases in private museum numbers at high population values (figure \ref{fig:p_pred_heatmap}, upper half), as such pattern does not correspond to that predicted by competition.
More in line with expectation is however the association between increases of population numbers and lower mortality at higher numbers of private museums (right half), as it indicates that audiences can facilitate museum survival.

@@

Museums which include the term "collections" in their name are the least likely to close (0.63 times the rate of museums with the name "museum", while those with the term foundation experience 23\% elevated closure rates (but not significantly higher).
However, institutions which include neither of these are the most likely to close, and have mortality rates almost three times as high (198\%) than institutions including the term "museum".
The usage of unconventional names might reflect adherence to museum standards which makes museums more understandable and familiar to third parties (e.g. audiences, other museums, companies or the government) and thereby facilitates interactions which are beneficial for long-term survival (e.g. attracting visitors, cooperations, sponsorships or subsidies).
However, as founders can choose the name of their museum arbitrarily, they might choose a name without reference to organizational forms associated with perpetuity precisely because their intention is not an institution enduring in perpetuity but rather a more temporally bounded project.
Some support for the latter interpretaton is given by the fact that two of the three museums for which \textcite{Velthuis_Gera_2024_fragility} identify a temporal intention as the closing reason do not self-identify as museum, collection or foundation.
A temporal intentions are however the exceptions among closing reasons identified by \textcite{Velthuis_Gera_2024_fragility} (p.6), and as similarly only a small fraction of private museum chooses such a name (140 museums, or 28\%), the overall proportion of these types of voluntary closures compared to all closures is presumably limited.
\subsection*{Discussion and Conclusion}

Previously the dozens of closures have been seen as indicator of their vulnerability and short-livedness \parencite{Adam_2020_close}, however this analysis shows the importance studying closing not in isolaton but in comparison to both non-closures and closures of comparable institutions.
While the booming numbers of private museums themselves have received substantial attention, less consideration has been given to the fact that large populations can generate dozens of closures without requiring high closing rates, especially when cloures are considered over a span of several years or decades.
In other words, the sheer size to which the private museum population has grown over the last decades has made it all but inevitable that some of them would close, and do so after only a few years.
Based on the current analysis, private museums will continue to close in the years to come; at hazard rates of around 1\% on average a handful of closures every year are to be expected.

Comparing private museums closures to closures of other organizational form advances our understanding of the sustainability of this new organizational form.
While private museums do close, the comparison to previous studies of museum closing rates provides no evidence that dependence on single founders constitutes a less stable organizational form than reliance on less centralized foundations, associations or corporate sponsorship.
Multiple reasons might be the at play that mitigate this risk of low diversification.
First, it might take more commitment to open a museum as an individual or couple compared to less involved art patronage of being a board member in existing institutions or donating works, which would result in private museums founders being more motivated to secure the continued existence of their institutions.
Furthermore, the centrality of the founder might leave little room to question who is responsible to provide additional funds in times of financial difficulties, whereas in organizations without a central figure each member may feel less responsible.
It might also be such commitment that leads founder to plan the future of their museum to such an extent that their death is not significantly associated with unvoluntary closing.

While so far little evidence points toward low diversification, i.e. reliance on the founder, as a key vulnerability of private museums to unvoluntary closures, the centrality of the founder may be more pronounced in voluntary closures. 
Indicated by unconventional naming choice, a subset of founders might from the very beginning envision the museum as temporary.
On the other hand, higher closing rates for unconventionally named museums might reflect absence of recognition and support by third parties in the environment.
A dimension where the influence of the environment can however be ascertained with greater certainty is the role of audiences and other private museums.
While the overall relationship is complex, for the majority of private museums increases in both variables are associated with higher closing rates.
Whereas it is plausible that the association between higher institution counts and closing rates reflects competitive pressures, it is less obvious what may lay behind observing a positive association behind larger population and closing risk. 
\subsubsection*{Limitations}

A number of limitations are worth considering:
So far only a small fraction (47 museums, 9.3\%) of museums has experienced the death of their founder, one of the primary reasons hypothesized to lead to the private museum closure. 
While so far these museums have not seen substantially elevated closing risks, these estimates will become more precise as more founders pass away.
As the currently large confidence interval (figure \ref{fig:p_surv_death}) for museums whose founders have died allows a wide range of survival chances, it is entirely possible that previous scholars have indeed correctly identified a key dependency (the current absence of evidence of it should not be taken as evidence of absence) and that therefore later analyses will a significantly association between founder death and closing risk as well as a higher average closing risk for private museums.\footnote{Similarly, as 80\% of private museums have been founded in the new millenium, only a few museums have so far reached an age above 30 years, which makes the estimation of the age-dependent hazard for this age range highly speculative.}
At the same time, it is also possible that private museums are generally less dependent on a single person (or couple) than has been assumed, and have in fact become institutionalized to an extent that most heirs of founders feel as motivated to continue the original founders' efforts as these founders themselves.

@@

The primary mechanism that has been argued to lead to museum closure is costs, and while this study measures it indirectly, primarily through indicators on the level of the founder, museum and environment, the financial situation of a museum is not used directly.
The primary reason for this is data availability.
While centralized non-profit financial information is available for American non-profits through the IRS/NCCS files, the much larger private museum population in the rest of the world is not subject to similar reporting requirements.
The effort to compile such information on a global scale ranges from extremely challenging (harmonizing different reporting standards, language barriers) to outright impossible for institutions who do not even (have to) report public financial information.
Similarly, other museum characteristics which were not considered here, such as as museum size, facilities, as well as marketing and outreach activities, might be related to long-term survival prospects.\footnote{While the original data collection included the assessment of various museum characteristics, such as floor and collection size, facilities, and cooperations, it became clear during data analysis that the data collected was unsuitable to construct accurate indicators from these findings due to difficulty in assessing them in many cases.
Especially for closed museums data collection proved difficult, as in many cases websites had been shut down, and while in many cases snapshots of the main website were available in the Wayback machine, less central pages (such as those with information about facilities and activities), were archived much less frequently, and sometimes inaccessible despite being archived due to changes in webtechnology (such as the retirement of Adobe flash).}
As such, private museums remain relatively intransparent, which does not prevent, but to some extent limits the identification of risk factors.
